---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document: 
    keep_md: true
---

```{r load-library}
# load necessary R packages
library(dplyr)

# set locale to prevent Chinese output
Sys.setlocale("LC_TIME", "C")
```

## Loading and preprocessing the data
```{r load-data}
# read data
data <- read.csv(unz("activity.zip", "activity.csv"), 
                 stringsAsFactors = FALSE)

# convert string to date
data$date <- as.Date(data$date)

# convert interval to minutes of a day
data$minutes <- floor(data$interval / 100) * 60 + data$interval %% 100
```

## What is mean total number of steps taken per day?

```{r step1}
# Calculate total steps per day
total_steps_per_day <- data %>% 
    group_by(date) %>% 
    summarize(steps_per_day = sum(steps, na.rm = TRUE))

# Plot histogram
hist(total_steps_per_day$steps_per_day, breaks = seq(0, 25000, 2500),
     xlab = "Total Steps per Day", xlim = c(0, 25000),
     main = "Histogram of Total Number of Steps Taken Each Day")
```
The mean and median total number of steps taken per day are `r mean(total_steps_per_day$steps_per_day)` and `r median(total_steps_per_day$steps_per_day)`, respectively.

## What is the average daily activity pattern?

```{r step2}

# Calculate average daily activity based on interval
daily_activity <- data %>% 
    group_by(minutes) %>% 
    summarize(activity = mean(steps, na.rm = TRUE))

# Plot daily activity pattern vs Time
plot(daily_activity$minutes, daily_activity$activity, 
    type = "l", xaxt = "n", xlab = "Hour",
    ylab = "Activity", main = "Daily Activity Pattern vs Time")
axis(1, at = seq(0, 1440, 60), labels = seq(0, 24, 1))
```

The `r which.max(daily_activity$activity)`-th item, i.e. the `r floor(data$interval[which.max(daily_activity$activity)] / 100)`:`r data$interval[which.max(daily_activity$activity)] %% 100` interval, contains the maximum number of steps.

## Imputing missing values

There are `r sum(is.na(data$steps))` missing values in the dataset. We choose to fill them with mean of activity of each time interval averaged across all days, round to nearest integer.

```{r step3}
# Create a new dataset
new_data <- data

# Fill new data set with the mean of activity of each time interval 
# averaged across all days.
new_data$steps[which(is.na(data$steps))] <- as.integer(round(daily_activity$activity[(data$minutes[which(is.na(data$steps))] + 5)/5]))

# Calculate total steps per day based on new dataset
new_total_steps_per_day <- new_data %>% 
    group_by(date) %>% 
    summarize(steps_per_day = sum(steps))

# Plot histogram
hist(new_total_steps_per_day$steps_per_day, breaks = seq(0, 25000, 2500),
     xlab = "Total Steps per Day", xlim = c(0, 25000),
     main = "New Histogram of Total Number of Steps Taken Each Day")
```

The mean and median total number of steps taken per day are `r mean(new_total_steps_per_day$steps_per_day)` and `r median(new_total_steps_per_day$steps_per_day)`, respectively. They are different from original values, i.e. mean `r mean(total_steps_per_day$steps_per_day)` and median `r median(total_steps_per_day$steps_per_day)`. The percentage of change of mean is `r (mean(new_total_steps_per_day$steps_per_day) - mean(total_steps_per_day$steps_per_day))/mean(total_steps_per_day$steps_per_day)*100`% which is larger than the percentage of change of the median `r (median(new_total_steps_per_day$steps_per_day) - median(total_steps_per_day$steps_per_day))/median(total_steps_per_day$steps_per_day)*100`%.

## Are there differences in activity patterns between weekdays and weekends?

```{r step4}
# Create a new factor variable in the dataset with two levels ¡V ¡§weekday¡¨
# and ¡§weekend¡¨ indicating whether a given date is a weekday or weekend
# day.
new_data$day_type <- 
    as.factor(ifelse(weekdays(new_data$date) %in% c("Sunday", "Saturday"), "weekend", "weekday"))

# Calculate daily activity averaged by interval and group by day_type
weekday_activity <- new_data %>% 
    group_by(day_type, minutes) %>%
    summarize(activity = mean(steps))

# Load lattice package
library(lattice)

# Plot weekday's daily activity pattern vs Time
with(weekday_activity,
    xyplot(activity ~ minutes | day_type, 
        layout = c(1, 2), type = "l", xaxt = "n", xlab = "Hour", 
        ylab = "Activity",
        scales=list(x=list(at=seq(0, 1440, 60), labels=seq(0, 24, 1)))
    )
)
```